---
title: "Steps assignement"
author: "Francesca Arcidiacono"
date: "21 ottobre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(ggplot2)
```

## Reading the dataset##

You need to get your wk and read the data through the read.csv function. Clean the data of the NA and show the first lines. 

```{r reading the data}
setwd("C:/Users/utente/Desktop/Learning/R Programming/Assign two months of steps")
data_steps <- read.csv("activity.csv")
data <- data_steps[ with (data_steps, { !(is.na(steps)) } ), ]
head(data)
```

##Creating an histogram with the total nb of steps per day and calculating mean and median##

Now you can find the total of the steps per day. You need to install the dplyr package. 

```{r creating the histogram}
by_day <- group_by(data, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day
hist(steps_by_day$total, main="Histogram with number of steps per day", 
xlab="Number of steps by day")
summary(steps_by_day)
```
The summary function on steps_by_day shows that the mean is 10766 and the median is 10765. 

##Time series plot of the average number of steps taken##

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r creating the plot}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
      main="Average number of steps over all days", xlab="Interval", 
      ylab="Average number of steps")
```


##Five minutes interval with max steps##

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r five min with max steps}
max_steps_row <- which.max(steps_by_interval$steps)
steps_by_interval[max_steps_row, ]
```

The 5 minutes interval with the maximum number of steps is the 835 and has 206.1698 steps.


##Code to describe and show a strategy for inputting missing data#

```{r inputting missing data}
sum(is.na(data_steps))
```
 The number of rows with NA si 2304

```{r inputting missing data2}
data_imputed <- data_steps
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
```


I've created new data set data_no_na which equals to data_row but without NA's. All NA's are replaced with mean of 5-minute interval

```{r inputting missing data3}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```

##Histogram of total number of steps taken each day after NA is imputed##

```{r hist}
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
mean(df_imputed_steps_by_day$steps)
median(df_imputed_steps_by_day$steps)
mean(steps_by_day$total)
median(steps_by_day$total)
```

Mean values stays the same but therer is slight difference in meadian value.



